---
title: "In-Class Activity:  Dates and Times"
author: ""
date: "January 14/15, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Setup

We are now using a new package called `lubridate`.  To install it, copy the following into your console

```
install.packages("lubridate")
```

You only need to run this once, ever.

Do NOT include this line in a code chunk - if you do, the package will be re-installed every single time you knit the file!

Then run the following chunk to load up the packages.

```{r}

library(tidyverse)
library(lubridate)
```


## Warm-up exercises

Answer the following using code (even if you know the answer already).

1. Save your birthday as a Datetime object in R.  
2. What day of the week does/did your birthday fall on in 2019?
3. On what day of the week should you plan your 100th birthday party?
4. Find the date that is exactly 9 months before your birthday.  If it falls on a holiday, think about the implications.
5. The following code loads a dataset giving the date ranges for the 12 Zodiac signs.  By making appropriate adjustments and calculations, find your Zodiac sign.

```{r}
zodiac <- read.csv("./Data/zodiac.csv")
```


## Solve the mystery

Speaking of the Zodiac, one of the most famous mysteries in California history is the identity of the so-called "Zodiac Killer", who murdered 7 people in Northern California between 1968 and 1969.

A new murder was committed last year in California, suspected to be the work of a new Zodiac Killer on the loose.  Fortunately, a camera was planted at the location where the body was found.  The FBI has provided you with a list of people who entered the building last year, and what date and time they entered, in Greenwich Mean Time (GMT).


```{r}
suspects <- read.csv("./Data/suspect_times.csv")
```

Unfortunately, the date and time of the murder is not known.  You have been hired to crack the case.  Use the clues below to discover the murderer's identity.

```{r}
ted <- suspects %>% filter(Name == " Ted Cruz") %>% mutate(
  Time.Spotted = force_tz(ymd_hms(Time.Spotted), "America/Los_Angeles")
)
am(ted$Time.Spotted)
wday(ted$Time.Spotted, label = TRUE)
month(ted$Time.Spotted)
```


1. Based on the sunburn patterns on the body, the murder was not committed in the morning.

```{r}
suspects <- suspects %>% mutate(
  Time.Spotted = force_tz(ymd_hms(Time.Spotted), "America/Los_Angeles")
  ) %>% filter(!am(Time.Spotted)) 

```


2. The room where the murder was committed is closed Tuesdays and Thursdays, so the murder did not happen on those days.

```{r}
suspects <- suspects %>% filter(!wday(Time.Spotted, label = TRUE) %in% c("Tuesday", "Thursday")) 
```


3. It is believed that the murder was committed within 5 weeks (35 days) of Thanksgiving, 2018. 

```{r}
tgiving <- make_date(year = 2018, month = 11, day = 22)
suspects <- suspects %>% filter(Time.Spotted >= tgiving + 35, Time.Spotted <= tgiving + 35) 
```










