---
title: "In-Class Activity:  12 Days of Christmas"
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Setup

The song "12 Days of Christmas", written around 1780, tells the tale of many gifts a person recieves in the days leading up to Christmas. [https://en.wikipedia.org/wiki/The_Twelve_Days_of_Christmas_(song)](https://en.wikipedia.org/wiki/The_Twelve_Days_of_Christmas_(song)) These gifts repeat; on the first day, the narrator recieves

```
A partridge in a pear tree.
```

On the last day, he or she recieves

```
Twelve Drummers Drumming
Eleven Pipers Piping
Ten Lords a Leaping
Nine Ladies Waiting
Eight Maids a Milking
Seven Swans a Swimming
Six Geese a Laying
Five Golden Rings
Four Calling Birds
Three French Hens
Two Turtle Doves
And a Partridge in a Pear Tree
```

Run the code below to load up a dataset called `xmas` that contains the crucial information about these gifts.  Your task will be to make R sing the whole song to you, using only this dataset.

```{r}

library(tidyverse)

xmas <- read.csv("https://www.dropbox.com/s/e584pryn8evm1gz/xmas.csv?dl=1")
```


## Step One:  Plurals

Note that the gifts are listed in singular: for example, on day 5 the narrator recieves "5 golden rings", but the entry for the gift on day 5 simply says "ring".  Write a function called "pluralize_gifts" that takes a gift and returns the appropriate plural. Some code structure is provided for you below.

*Hint: Note that the gifts on days 6 and 9 have unusual pluralizations.*
*Hint:  The functions `paste()` and `paste0()` should be useful, as well as the stringr functions from last week.*

Important:  You should **absolutely not** hard-code anything into this function.  For example, the word "rings" should not appear in the function.  I should be able to give it any list of gifts and any list of days, and get back the plurals.

```{r}

pluralize_gifts <- function(gift, day){
  
  skip <- ## which gifts do we not simply put an "s" at the end of?
  
  gifts <- gifts %>%  ## make the special adjustments
        
  gifts[!skip] <- ## put and s at the end of the rest
  
  return(gifts)
  
} 

```



Now, use `mutate` with your new function to add a variable called `Plural.Gifts` to the `xmas` dataset.


## Step Two: Creating sentences


Write a function called `make_phrase` that takes as input the necessary information, and returns a phrase.  For example, 

`make_phrase(num = 10, item = "lords", verb = "a-leaping", adjective = "", location = "")`

should return

```
"10 lords a-leaping"
```

**Hint: the function `str_replace_na` may be useful.**

```{r}
make_phrase <- function(num, item, verb, adjective, location){
  
 
  
  return(phrase)
}


```

Use this function to make a new variable called `phrases` that contains all the gifts.


## Step Three:  A for-loop

Use a for loop to sing the song.  At each step, your loop should:
- Print out an introductory phrase, such as  "On the first day of Christmas, my true love sent to me"
- Print out all the necessary gift phrases, with correct punctuation between them.

*Hint: Use `print()` make R show the sentences/phrases.*
*Hint: You  may want a secondary loop!*

```{r}

for (i in 1:12){
  
  
  print(paste("On the ", xmas$Day.in.Words[i], "day of Christmas, my true love sent to me"))
  
  
}


```

## Step Four:  Using apply and map.

Write a function called `sing_line` that takes as input a number corresponding to a row of the dataset `xmas` and returns the entire section of the song for that day of Christmas.  For example, 

`sing_line(xmas, 2)`

should return

```
On the second day of Christmas, my true love sent to me,
two turtle doves and a partridge in a pear tree.
```

*Hint: This should be a quick copy-paste, since your function will basically perform all the tasks inside your for loop!*

Then have R sing me the Twelve Days of Christmas song using (a) an apply function and (b) a map function.


## Step Five:  One function to rule them all.

Combine steps 1-4 into one function called `sing_song`.  That is, this function should take as input:
- the vector of days
- the vector of (unpluralized) gifts
- etc

*Note: You are encouraged to make smaller functions - such as the ones you used earlier to pluralize gifts or write phrases - and reference these inside your primary function.*

Test your function on the following code, which should return the whole "12 Days of Christmas" song:

```{r}
xmas <- read.csv("https://www.dropbox.com/s/e584pryn8evm1gz/xmas.csv?dl=1")

sing_song(xmas)
```

When you have finished, open the file "7_FP_Assignment", and copy your function into the code chunk, follow the directions, and knit.  Turn this in to PolyLearn.




