---
title: "In-Class Activity:  12 Days of Christmas"
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

The song "12 Days of Christmas", written around 1780, tells the tale of many gifts a person recieves in the days leading up to Christmas. [https://en.wikipedia.org/wiki/The_Twelve_Days_of_Christmas_(song)](https://en.wikipedia.org/wiki/The_Twelve_Days_of_Christmas_(song)) These gifts repeat; on the first day, the narrator recieves

```
A partridge in a pear tree.
```

On the last day, he or she recieves

```
Twelve Drummers Drumming
Eleven Pipers Piping
Ten Lords a Leaping
Nine Ladies Waiting
Eight Maids a Milking
Seven Swans a Swimming
Six Geese a Laying
Five Golden Rings
Four Calling Birds
Three French Hens
Two Turtle Doves
And a Partridge in a Pear Tree
```

Run the code below to load up a dataset called `xmas` that contains the crucial information about these gifts.  Your task will be to make R sing the whole song to you, using only this dataset.

```{r}

library(tidyverse)

xmas <- read.csv("https://www.dropbox.com/s/e584pryn8evm1gz/xmas.csv?dl=1")
```


## Step One:  Plurals

Note that the gifts are listed in singular: for example, on day 5 the narrator recieves "5 golden rings", but the entry for the gift on day 5 simply says "ring".  Write a function called "pluralize_gifts" that takes a gift and returns the appropriate plural. Some code structure is provided for you below.

*Hint: Note that the gifts on days 6 and 9 have unusual pluralizations.*
*Hint:  The functions `paste()` and `paste0()` should be useful, as well as the stringr functions from last week.*

Important:  You should **absolutely not** hard-code anything into this function.  For example, the word "rings" should not appear in the function.  I should be able to give it any list of gifts and any list of days, and get back the plurals.

```{r}

pluralize_gifts <- function(gifts, days){
  
  skip <- days == 1 | str_detect(gifts, "y$") | str_detect(gifts, "oo")
  
  gifts <- gifts %>%
    str_replace("oo", "ee") %>%
    str_replace("y$", "ies")
        
  gifts[!skip] <- paste0(gifts[!skip], "s")
  
  return(gifts)
  
} 

```



Now, use `mutate` with your new function to add a variable called `Plural.Gifts` to the `xmas` dataset.

```{r}
xmas <- xmas %>%
  mutate(
    Plural.Gifts = pluralize_gifts(Gift.Item, Day)
  )
```



## Step Two: Creating sentences

Write a function called `make_phrase` that takes as input the necessary information, and returns a phrase.  For example, 

`make_phrase(num = 10, item = "lords", verb = "a-leaping", adjective = "", location = "")`

should return

```
"10 lords a-leaping"
```

```{r}
make_phrase <- function(num, item, verb, adjective, location){
  
   verb <- str_replace_na(verb, "")
  adjective <- str_replace_na(adjective, "")
  location <- str_replace_na(location, "")
  
  phrase <- paste(num, adjective, item, verb, location)
  phrase <- phrase %>% 
    str_replace_all("\\s+", " ") %>%  # in case two spaces together
    str_replace_all("(^| )1 ", " a ") %>%  # for single items
    str_trim()
  
  return(phrase)
  
}


```


Use this function to make a new variable called `phrases` that contains all the gifts.

```{r}
xmas <- xmas %>% mutate(
  phrases = make_phrase(Day, Plural.Gifts, Verb, Adjective, Location)
)
```

## Step Three:  A for-loop

Use a for loop to sing the song.  At each step, your loop should:
- Print out an introductory phrase, such as  "On the first day of Christmas, my true love sent to me"
- Print out all the necessary gift phrases, with correct punctuation between them.

*Hint: You  may want a secondary loop!*

```{r}

for (i in 1:12){
  
  my_line <- paste("On the", xmas$Day.in.Words[i], "day of Christmas, my true love sent to me")
  
  if(i == 1){
    
    my_line <- paste(my_line, xmas$phrase[i])
    
  } else{
  
    for (j in i:2){
    
      my_line <- paste0(my_line, " ", xmas$phrase[j], ", ")
    
    }
    my_line <- paste(my_line, "and", xmas$phrase[1])
  }
  
  
  print(my_line)

}


```

## Step Four:  Using apply and map.

Write a function called `sing_line` that takes as input number corresponding to a row of the dataset `xmas` and returns the entire section of the song for that day of Christmas.  For example, 

`sing_line(xmas, 2)`

should return

```
On the second day of Christmas, my true love sent to me,
two turtle doves 
and a partridge in a pear tree.
```

*Hint: This should be a quick copy-paste, since your function will basically perform all the tasks inside your for loop!*

Then have R sing me the Twelve Days of Christmas song using (a) an apply function and (b) a map function.

```{r}
sing_line <- function(dataset, i){
  
   my_line <- paste("On the", dataset$Day.in.Words[i], "day of Christmas, my true love sent to me")
  
  if(i == 1){
    
    my_line <- paste(my_line, dataset$phrase[i])
    
  } else{
  
    for (j in i:2){
    
      my_line <- paste0(" ", my_line, dataset$phrase[j], ", ")
    
    }
    my_line <- paste(my_line, "and", dataset$phrase[1])
  }
  
  
  return(my_line)
  
}


```

```{r}
sapply(1:12, function(x) sing_line(xmas, x))

map_chr(1:12, ~sing_line(xmas, .x))
```




## Step Five:  One function to rule them all.

Combine steps 1-4 into one function called `sing_song`.  That is, this function should take as input:
- the vector of days
- the vector of (unpluralized) gifts
- etc

And return the full song.

```{r}

sing_song <- function(dataset){
  
  dataset <- dataset %>%
    mutate(
      Plural.Gifts = pluralize_gifts(Gift.Item, Day),
      phrase = make_phrase(Day, Plural.Gifts, Verb, Adjective, Location)
    )
  
  map_chr(1:12, ~sing_line(dataset, .x))
  
}

```



*Note: You are encouraged to make smaller functions - such as the ones you used earlier to pluralize gifts or write phrases - and reference these inside your primary function.*

Test your function on the following code, which should return the whole "12 Days of Christmas" song:

```{r, echo = TRUE}
xmas <- read.csv("https://www.dropbox.com/s/e584pryn8evm1gz/xmas.csv?dl=1")

sing_song(xmas)
```

When you have finished, open the file "7_FP_Assignment", and copy your function into the code chunk, follow the directions, and knit.  Turn this in to PolyLearn.




