---
title: "8_Activity_Testing"
author: ""
date: "February 6, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(broom)

data("starwars")
```

# Instructions

The dataset "starwars" has been loaded for you.  Take a look at it and understand the variables. 

## Writing your own t-test

Write a function called `run_t_test` that takes as input two vectors of numbers, and an option for the alternative hypothesis, and gives back the results of a t-test for equality of means.  You may not use the function `t.test()` except to check that your version is working correctly.

```{r}
run_t_test <- function(x, y, alternative = "two.sided"){
  
  x <- x[!is.na(x)]
  y <- y[!is.na(y)]
  
  mx <- mean(x)
  my <- mean(y)
  
  sdx <- sd(x)
  sdy <- sd(y)
  
  nx <- length(x)
  ny <- length(y)
  
  se <- sqrt(sdx^2/nx + sdy^2/ny)
  
  t_score <- (mx - my)/se
  
  dof <- min(nx, ny) - 1
  
  if (alternative == "two.sided"){
    
    pval <- 2*pt(-abs(t_score), df = dof)
    
  } else if (alternative == "less") {

    pval <- pt(t_score, df = dof)
    
  } else if (alternative == "greater") {
    
    pval <- 1-pt(t_score, df = dof)
    
  }
  
  
  
  return(list(test.stat = t_score, df = dof, p.val = pval))
}
```

Use your t-test function to answer the following:

1. Are Humans or Droids taller?

```{r}
h_heights <- starwars %>%
  filter(species == "Human") %>%
  pull(height)

d_heights <- starwars %>%
  filter(species == "Droid") %>%
  pull(height)


run_t_test(h_heights, d_heights)

```



2. Do men or women weigh more?

```{r}
m_weights <- starwars %>%
  filter(gender == "male") %>%
  pull(mass)

w_weights <- starwars %>%
  filter(gender == "female") %>%
  pull(mass)

run_t_test(m_weights, w_weights)
```


## Chi-square tests

Use the function `chisq.test()` to answer the following:

1. Is there a relationship between hair color and eye color?

```{r}
starwars <- starwars %>% mutate(
  Hair = case_when(
    hair_color == "brown" | hair_color == "black" ~ "dark_hair",
    hair_color == "none" ~ "bald",
    TRUE ~ "light_hair"
  ),
  Eyes = case_when(
    eye_color == "black" | eye_color == "dark" ~ "black",
    eye_color == "brown" ~ "brown",
    eye_color == "blue" ~ "blue",
    TRUE ~ "other"
  )
)

starwars %>% 
  count(Hair, Eyes) %>%
  spread(key = Hair, value = n) %>%
  select(-Eyes) %>%
  chisq.test() %>%
  tidy()
```



2. Do Droids tend to come from some homeworlds more than others?

```{r}
has_droids <- starwars %>% 
  filter(species == "Droid") %>%
  count(homeworld) %>%
  filter(n > 0) %>%
  pull(homeworld)

starwars %>% 
  filter(homeworld %in% has_droids) %>%
  filter(!is.na(species)) %>%
  mutate(
    species2 = species == "Droid"
  ) %>%
  count(species2, homeworld) %>%
  spread(species2, n) %>%
  select(-homeworld) %>%
  chisq.test() %>%
  tidy()

```



## Many tests (optional)

On which worlds is the height difference between men and women most pronounced?

```{r}
#### I studied male vs. non-male, but I asked the question wrong on the activity.  My fault!  If you studied male vs. female, only Tattoine has enough observations to do a t-test.


starwars <- starwars %>% 
  mutate(
    gender2 = case_when(
      gender == "male" ~ "male",
      TRUE ~ "non-male"
    )
  )
  
good_worlds <- starwars %>%
  count(homeworld, gender2) %>%
  spread(key= gender2, value = n) %>%
  filter(`non-male` >= 3, male >=3) %>%
  pull(homeworld)

dat <- starwars %>% filter(homeworld %in% good_worlds) %>% group_by(homeworld) %>% nest()

map_df(dat$data, ~tidy(t.test(height ~ gender2, data = .x)))
```

 



