---
title: "Bootstrapping"
author: Kelly Bodwin
date: 
output: 
  ioslides_presentation:
    fontsize: 10pt
    incremental: true
  # beamer_presentation:
  #   theme: "boxes"
  #   colortheme: "beaver"
  #   fonttheme: "structurebold"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
titanic <- read.csv("https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Datasets/titanic.csv")
library(tidyverse)
library(lubridate)
library(modelr)
library(broom)

data(mtcars)
```


```{r}
automatic_cars <- mtcars %>% 
  filter(am == 1)

manual_cars <- mtcars %>% 
  filter(am == 0)

manual_cars$hp
```

```{r}
set.seed(35)
my_automatic_boots <- replicate(1000, sample(automatic_cars$hp, replace=TRUE), simplify = FALSE)
my_automatic_boot_means <- map_dbl(my_automatic_boots, mean)

my_manual_boots <- replicate(1000, sample(manual_cars$hp, replace=TRUE), simplify = FALSE)
my_manual_boot_means <- map_dbl(my_manual_boots, mean)

diff_means <- my_automatic_boot_means - my_manual_boot_means

quantile(diff_means, c(0.025, .975))
```


```{r}

boot_CIs <- data.frame()

diff_means <- mean(automatic_cars$hp) - mean(manual_cars$hp)

boot_it(automatic_cars$hp)

boot_it(manual_cars$hp)
```

```{r}
my_sample <- rbinom(10, 1, .8)

boot_it(my_sample)
```





# Estimation and Confidence


## Estimating the mean

Suppose we want to find the mean of a population, but we only have one sample, of size 100.  Our best guess is

```{r}
my_sample <- rnorm(100)
mean(my_sample)
```

Not bad!  (The true mean is 0.)  But how confident are we?

## Confidence

```{r}
my_test <- tidy(t.test(my_sample))
my_test
```

## Confidence

We are 95% confident that the true mean is between `r my_test$conf.low` and `r my_test$conf.high`.


BUT!  What assumptions did this depend on?

## Normality

```{r}
ggplot(, aes(x = my_sample)) + 
  geom_histogram(aes(y = ..density..), bins = 10, fill = "cornflowerblue")+
  stat_function(fun = dnorm, color = "darkred", size = 2) 
```


## Estimating the mean with skew

```{r}
skewed_sample <- rexp(1000)
```

```{r, echo = FALSE, warning = FALSE}
ggplot(, aes(x = skewed_sample)) + 
  geom_histogram(aes(y = ..density..), bins = 10, fill = "cornflowerblue")+
  stat_function(fun = dnorm, args = list(mean = mean(skewed_sample), sd = sd(skewed_sample)), color = "darkred", size = 2) +
  xlim(c(-5, 5))
```

## Confidence

```{r}
my_test_2 <- tidy(t.test(skewed_sample))
my_test_2
```

## Ask yourself...

* Do you trust the t-confidence interval?
* Is the mean the best measure of center?
* How would you make a confidence interval for the median?

# Pull yourself up by your bootstraps!

## Idea:

* A large enough representative sample from a population is a good approximation for the distribution of the population.
* We could do better inference we had more samples.

Thus:  Re-sample from our sample!

## Bootstrap it:

```{r}
sample(skewed_sample, replace = TRUE)
# why is replace = TRUE important?
```

## Bootstrap it:

```{r}
my_boots <- replicate(1000, sample(skewed_sample, replace = TRUE), simplify = FALSE)
```

```{r, echo = FALSE, warning = FALSE}
ggplot(, aes(x = unlist(my_boots))) + 
  geom_histogram(aes(y = ..density..), bins = 10, fill = "cornflowerblue")+
  stat_function(fun = dnorm, args = list(mean = mean(skewed_sample), sd = sd(skewed_sample)), color = "darkred", size = 2) +
  xlim(c(-5, 5)) 
```

## Bootstrapped confidence

```{r}
my_stats <- data.frame(
  means = map_dbl(my_boots, mean),
  medians = map_dbl(my_boots, median)
)
head(my_stats)
```

## Plot it

```{r}
my_stats %>% 
  gather("Stat", "Value") %>%
ggplot(aes(x = Value, fill = Stat)) +
  geom_histogram(bins = 20)
```

## Confidence intervals

What range catches 95% of our bootstrap estimates?

```{r}
my_test_2 %>% select(conf.low, conf.high)
quantile(my_stats$means, c(0.025, 0.975))
quantile(my_stats$medians, c(0.025, 0.975))
```

## Ask yourself...

* Do you trust the t-confidence interval?
* Is the mean the best measure of center?
* How would you make a confidence interval for the median?

# Very small samples

## Bootstrap vs. t-test

```{r}
small_sample <- rbinom(10, 1, 0.8)
tidy(t.test(small_sample)) %>% select(conf.low, conf.high)
```

## Bootstrap vs. t-test

```{r}
my_boots <- replicate(1000, sample(small_sample, replace = TRUE), simplify = FALSE)
boot_means <- map_dbl(my_boots, mean)
quantile(boot_means, c(0.025, .975))
```

## Bootstrap vs. t

```{r, echo = FALSE}
boot_it <- function(sample){
  
  my_boots <- replicate(1000, sample(sample, replace = TRUE), simplify = FALSE)
  boot_means <- map_dbl(my_boots, mean)
  return(quantile(boot_means, c(0.025, .975)))
  
}
```

```{r}
boot_CIs <- data.frame()
t_CIs <- data.frame()
means <- c()
set.seed(35)
for(i in 1:10){
  
  my_sample <- rbinom(10, 1, .8)
  
  means <- c(means, mean(my_sample))
  
  boot_CIs <- rbind(boot_CIs, boot_it(my_sample))
  
  my_test <- tidy(t.test(my_sample)) %>% select(conf.low, conf.high)
  
  t_CIs <- rbind(t_CIs, my_test)
  
}
```

What is happening in this code?

```{r, echo = FALSE}
names(boot_CIs) <- c("lower", "upper")
names(t_CIs) <- c("lower", "upper")
final_results <- full_join(boot_CIs, t_CIs) %>%
  mutate(
    try = rep(1:10, 2),
    type = c(rep("Bootstrap", 10), rep("t", 10)),
    means = rep(means, 2)
  )
  
```



## Bootstrap vs. t

```{r, echo = FALSE}
final_results %>% arrange(try) %>%
ggplot(aes(x = means, fill = type)) + 
    geom_errorbarh(aes(xmin=lower, xmax=upper, y = 1:20, color = type), size = 2) +
  geom_point(aes(x = means, y = 1:20))
```


# Activity

## Automatic versus manual cars

Load the `mtcars` dataset into R by typing `data(mtcars)`.

This dataset contains information about 32 different cars.

Use bootstrapping to make a 95\% confidence interval for the difference in horsepower (`hp`) between automatic (`am = 0`) and manual (`am = 1`) cars.

Write a sentence summarizing your findings in the PolyLearn assignment.