---
title: "Exploring Data with the Tidyverse"
author: Kelly Bodwin
date: 
output: html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.align = 'center', fig.path='Figs/',
                      echo=TRUE, warning=FALSE, message=FALSE,eval=TRUE)
```

# Examine the Data

## The Dataset

- Information about passengers on the HMS Titanic

```{r, include = FALSE}
library(tidyverse)
titanic = read.csv('https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Datasets/titanic.csv')
```

```{r}
head(titanic)
```

## What do we know?

```{r}
str(titanic)
```


- How many passengers were there?
- What information did we collect about them?


```{r}
titanic[118,]
```

Tell me about this person!


# Get a feel for the variables

## How many passengers survived?

```{r}
titanic %>% count(Survived)
```

- Note that this information was only displayed, not saved!

- If you wanted to store your answer for later use...

```{r}
# base R
num_survived <- sum(titanic$Survived == 1)

# Tidyverse
num_survived <-  titanic %>% 
  tally(Survived == 1) %>% 
  pull(n)
## try it without the "pull(n)" step.  What is the difference?

```


```{r, echo = FALSE}
titanic %>%
  ggplot(aes(x = Survived)) + geom_bar()
```

- Yikes, that's ugly!

- Goal: Recode Survived to be a factor with values "Lived", "Died".

```{r}
titanic <- titanic %>%   # we are rewriting "titanic"
  mutate(
    Survived = factor(Survived, 
                      levels = c(0,1), # old values
                      labels = c("Died", "Lived") # new values
    )
  )

# by naming the new variable "Survived", we overwrite the old one

```

- Check yourself: Did that work?

```{r}
titanic %>% count(Survived)
```

Make a better plot:

```{r, echo = FALSE}

```


## Your Turn

- How many passengers were in each travel class?
- How many of each gender?
- How many of each gender *within* each travel class?  Make a plot.


# Research question: Looking at specific individuals.

## Did the person with the most expensive ticket survive?


```{r}
titanic %>% top_n(1, Fare)
# arrange(Fare) is okay too!
# But it prints ALL the answers

```

- Why did this give us 3 answers instead of 1 like we asked?


## Did the youngest person in third class survive?

```{r}
titanic %>% filter(Pclass == 3) %>% top_n(1, desc(Age))

```

## Your turn

- Did the richest woman in second class above the age of 30 survive?


# Research question: What explains why people lived or died?

## Gender

- Were men more likely to die than women?

```{r}
titanic %>% count(Sex, Survived)
```

- Hmmmm.... maybe.  
- (Bonus: what statistical test would you use for this question?)


- What percent of passengers survived in each group (men vs women)?

```{r}
titanic %>% 
  group_by(Sex) %>%
  summarize(pct_survived = mean(Survived == "Lived"))

```
- (Think:  Why did "mean" work?)
- (Try it:  What does the output look like if you remove the last line and pipe?)


- We can also visualize the data to answer our question.  Which plot will address the question better?


```{r}

titanic %>% 
  ggplot(aes(x = Survived, fill = Sex)) + geom_bar()

titanic %>% 
  ggplot(aes(x = Sex, fill = Survived)) + geom_bar()

titanic %>% 
  ggplot(aes(x = Survived)) + geom_bar() + facet_wrap(~Sex)

titanic %>% 
  ggplot(aes(x = Sex)) + geom_bar() + facet_wrap(~Survived)
```


- How can we best improve this plot to strengthen our point?


```{r}


titanic %>% 
  ggplot(aes(x = Survived, fill = Sex)) + geom_bar(position = "fill")

titanic %>% 
  ggplot(aes(x = Survived, fill = Sex)) + geom_bar(position = "dodge")


```

- What do you conclude?

## Your Turn

- Was one travel class more likely to survive than another?
- What is the relationship between gender, travel class, and survival?


# Extra practice

## Your turn

Try to use the same thought process with boxplots to answer these questions:

- Is there a price difference across the three classes?
- Is there an age difference across the three classes?


# Complex Research Question: What can we say about family, wealth, and survival?

## My thought process:

I'm not sure why we count siblings/spouses and parents/children separately.  Let's make new variables with clearer measurements.

```{r}
titanic <- titanic %>% mutate(
  Num.Family = Siblings.Spouses.Aboard + Parents.Children.Aboard,
  Has.Family = Num.Family > 0
)
```


Check that my variables worked

```{r}
titanic %>% 
  select(Num.Family, Has.Family) %>% 
  summary()
```


Wealth can be measured in two ways:  Class, or Fare.  I would like to come up with categories based on these.

First, I want to know what the spread of Fare is.

```{r}
summary(titanic$Fare)
```

Alright, so most people paid less thatn \$31, and some paid less than \$8.

I think I will call people "Rich" if they are in 1st class or paid over \$100, and "Poor" if they are in third class or paid less than \$8.

```{r}
titanic <- titanic %>% mutate(
  wealth = case_when(
    Pclass == 3 | Fare < 8 ~ "Poor",
    Pclass == 1 | Fare > 100 ~ "Rich",
    TRUE ~ "Medium"
  ),
  wealth = factor(wealth, levels = c("Poor", "Medium", "Rich")) # reorder
)
```


Let me double check that my definitions split the population reasonably.

```{r}
titanic %>% count(wealth)
```

Great, seems okay!

Now I need a plot.  I think I would like to compare the number of family members for the various wealth levels, divided by whether they lived or died.

```{r, echo = FALSE}

titanic %>% 
  ggplot(aes(x = wealth, y = Num.Family, fill = Survived)) + geom_boxplot()

```

Hmmmm not so great.  Perhaps simply if they had family or not would make more sense.


```{r, echo = FALSE}

titanic %>% 
  ggplot(aes(x = Has.Family, fill = Survived)) + geom_bar(position = "fill") + facet_wrap(~wealth)

```


What do you conclude?

## YOUR TURN

Can you come up with different ways to address this question?