---
title: "Tidy Data; gathering and spreading"
author: Kelly Bodwin
date: 
output: 
  ioslides_presentation:
    fontsize: 10pt
    incremental: true
  # beamer_presentation:
  #   theme: "boxes"
  #   colortheme: "beaver"
  #   fonttheme: "structurebold"
---

```{r, include = FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.align = 'center', fig.path='Figs/',
                      echo=TRUE, warning=FALSE, message=FALSE,eval=TRUE)

library(tidyverse)

hiphop <- read.csv("https://www.dropbox.com/s/1qqyshx5ikt9zoc/hiphop.csv?dl=1")
```

# Tidy data, and grouping

## A toy example dataset

```{r, echo = FALSE}
set.seed(191)
dat <- data.frame(group = rep(c("a", "b", "c"), 3), meas1 = runif(9), meas2 = runif(9))
dat
```

## Tidy data

- The dataset (data frame) is ...

## Tidy data

- The dataset is `dat`
- The variables are ...

## Tidy data

- The dataset is `dat`
- The variables are `group`, `meas1`, `meas2`
- Observations consist of ...

## Tidy data

- The dataset is `dat`
- The variables are `group`, `meas1`, `meas2`
- Observations consist of two measurements for each group.  There are three such observations in the data for each group.

## Group_by makes new datasets (sort of)

```{r}
dat %>% top_n(2, meas1)

dat %>% group_by(group) %>% top_n(2, meas2)
```

## Summarize "squishes" these datasets

- But we have to tell it how to combine all the observations of `meas`.  Mean? Median? Sum? Max?
- All variables we can't "squish" get dropped
```{r}

dat %>% group_by(group) %>% 
  summarize(avg_meas = mean(meas1)) %>%
  top_n(1, avg_meas)

```
## To squish more than one variable:  _all

```{r, error = TRUE}
dat %>% group_by(group) %>% 
  summarize_all(mean)
```

## To squish more than one variable:   _if

```{r}
dat %>% group_by(group) %>% 
  summarize_if(is.numeric, funs(mean, median))
```

## To squish more than one variable:  _at 

```{r}
dat %>% group_by(group) %>% 
  summarize_at(vars(meas1, meas2), funs(mean, median))
```

## Group_by versus distinct

```{r}
dat %>% group_by(group) %>% summarize(avg_meas = mean(meas1))
dat %>% distinct(group, .keep_all = TRUE)
```


# Spread

# Question: What words are more preferred by hiphop fans than nonfans?

## On the lab assignment, used full_join
```{r}
fans <- hiphop %>% 
  filter(hiphop > 7) %>%
  group_by(word) %>% 
  summarize(avg_fam = mean(familiarity))

not_fans <- hiphop %>% 
  filter(hiphop < 4) %>%
  group_by(word) %>% 
  summarize(avg_fam = mean(familiarity))

full_join(fans, not_fans)
```

## Automatically renames columns if "by" is specified

```{r}
full_join(fans, not_fans, by = "word")
```


## Let's see the answer!

```{r}
new <- full_join(fans, not_fans, by = "word") %>%
  mutate(
    diff = avg_fam.x - avg_fam.y
  ) 

new %>% arrange(diff)
```


## But, a more elegant approach:
```{r}
fans <- hiphop %>% 
  mutate(
    is_fan = case_when(
      hiphop > 7 ~ "Fan",
      hiphop < 3 ~ "Not Fan",
      TRUE ~ "Sorta Fan"
    ),
    is_fan = factor(is_fan, levels = c("Fan", "Sorta Fan", "Not Fan"))
  ) %>% 
  select(word, familiarity, is_fan, hiphop, age)
 
head(fans)
```

## This is nice for plotting...
```{r}
fans %>% 
  ggplot(aes(x = is_fan, y = age, fill = is_fan)) + geom_boxplot()
```

## And for grouping and summarizing...
```{r}
fans %>% 
  group_by(is_fan) %>%
  summarize(avg_age = mean(age))
```

## But what about making a new variable?
```{r}
head(fans)
```

## But what about making a new variable?
```{r, eval = FALSE}
fans %>% mutate(
  diff = Fan - `Not Fan`
)

```

## What went wrong?
```{r, error = TRUE}
fans %>% mutate(
  diff = Fan - `Not Fan`
)

```

## There is no way to do this by subtracting variables.
```{r, eval = FALSE}
fans %>% mutate(
  diff = ....?
)

```

## (Not even in Base R)
```{r, eval = FALSE}
fans$diff <- 1

yes_fan <- fans$is_fan == "Fan"
fans$diff[yes_fan] <-  ...?
```

## Spread, to the rescue!

```{r}
fans %>% 
  group_by(is_fan, word) %>%
  summarize(avg_fam = mean(familiarity)) %>% 
  ### <b>
  spread(key = is_fan, value = avg_fam) 
### </b>
 
```

## Spread

- `key` is the *categorical variable* (i.e., a *factor*) that is going to convert to variable names
- `value` is any variable, usually quantitative, that is measured for each *key* category

## Now our mutate() works!
```{r}
fans <- fans %>% 
  group_by(is_fan, word) %>%
  summarize(avg_fam = mean(familiarity)) %>% 
  spread(key = is_fan, value = avg_fam) %>%
  ### <b>
  mutate(
    diff = Fan - `Not Fan`
  )### </b>
```

## Final answer:
```{r}
fans %>% top_n(1, diff)
fans %>% top_n(1, desc(diff))
```

# Gather

# Question:  Do men and women prefer different genres of music?

## Let's narrow down the data a bit

```{r}

genres <- hiphop %>% 
  distinct(subj, .keep_all = TRUE) %>%
  group_by(sex) %>%
  summarize_at(
    vars(classical, folk, rock, country, pop, alternative, hiphop), 
    funs(mean))
  
```

## Plotting

- I'd like to plot the 7 genres scores, for men and for women.
- But how???

```{r, eval = FALSE}
genres %>% ggplot(fill = sex, y = ...?, x = ...?)
```

## Gather to the rescue!
```{r}
genres %>% gather()
```

## Let's give the columns better names
```{r}
genres %>% gather(key = genre, value = avg_score)
```

## Gather

- `key` is the name you want to give the *categorical variable* whose values will be the former variable names
- `value` is name you want to give the variable of collapsed values for all categories

## Keep the column "sex" from being gathered
```{r}
genres %>% gather(key = genre, value = avg_score, -sex)
```

## Now plotting is easy!
```{r}
genres %>% gather(key = genre, value = avg_score, -sex) %>%
  ggplot(aes(x = genre, y = avg_score, fill = sex)) + geom_col(position = "dodge")
```

# Your Turn

## Re-create the following plot.  

```{r, echo = FALSE}
hiphop %>% 
  filter(ethnic != "black") %>%
  distinct(subj, .keep_all = TRUE) %>%
  mutate(
    has_black_friends = blackMove > 0
  ) %>%
  select(jayz, barkley, boondocks, monique, bieber, has_black_friends) %>%
  gather(key = culture_question, value = score, -has_black_friends) %>%
  group_by(culture_question, has_black_friends) %>%
  summarize(avg_score = mean(score)) %>% 
  ungroup() %>%
  ggplot(aes(x = culture_question, y = avg_score, fill = has_black_friends)) + geom_col(position = "dodge") +
  ggtitle("Do non-black people with black friends know more about black culture?")
```

Turn in an html file with your source code and final plot for your in-class assignment.


