---
title: "Functional Programming"
author: Kelly Bodwin
date: 
output: 
  ioslides_presentation:
    fontsize: 10pt
    incremental: true
  # beamer_presentation:
  #   theme: "boxes"
  #   colortheme: "beaver"
  #   fonttheme: "structurebold"
---

# The general idea

```{r, include = FALSE}
titanic <- read.csv("C:/Users/kbodwin/Dropbox/Teaching/331/Data/titanic.csv")
library(tidyverse)
library(lubridate)
```

## The general idea

- What if you want to do something a whole bunch of times, or to every item in a list?  
- We want to be as efficient as possible!

```{r}
print("99 bottles of pop on the wall, 
      99 bottles of pop.  
      Take one down, pass it around, 
      98 bottles of pop on the wall.")
print("98 bottles of pop on the wall, 
      98 bottles of pop.  
      Take one down, pass it around, 
      97 bottles of pop on the wall.")

###.... and so on...
```

## Evoution 1: Make R do the math

```{r}
i <- 99
print(paste(i, "bottles of pop on the wall,", 
            i, "bottles of pop.  Take one down, pass it around,", 
            i - 1, "bottles of pop on the wall."))
i <- i - 1
print(paste(i, "bottles of pop on the wall,", 
            i, "bottles of pop.  Take one down, pass it around,", 
            i - 1, "bottles of pop on the wall."))

###.... and so on...
```

## Evolution 2: Loop it!

```{r}
for(i in 99:95){
  
  print( 
    paste(i, "bottles of pop on the wall,", 
          i, "bottles of pop.  Take one down, pass it around,", 
          i - 1, "bottles of pop on the wall.")
  )

  
}
```

## Evolution 3: Write a function.

```{r}
pop_song <- function(i){
  
 paste(i, "bottles of pop on the wall,", 
            i, "bottles of pop.  Take one down, pass it around,", 
            i - 1, "bottles of pop on the wall.")
  
}

for(i in 99:95){
  
 print(pop_song(i))
  
}
```

## Evolution 4: Apply functions

```{r}

sapply(99:95, pop_song)

```

## The apply functions

- Pay attention to what type of input they *take* and what type of output they *give*
- `sapply` takes a vector, returns a vector
- `lapply` takes a list, returns a list
- `apply` takes a dataset and a dimension (1 = row, 2 = col), returns a vector or list

## Evolution 5: Tidyverse


```{r}

map_chr(99:95, pop_song)

```

## What if I want some options in my function?

```{r}
pop_song <- function(i, drink){
  
 paste(i, "bottles of", drink, "on the wall,", 
       i, "bottles of ", drink, ".  Take one down, pass it around,",
       i-1, "bottles of", drink, "on the wall.")
  
}
```

## Apply has trouble!

```{r, error = TRUE}

sapply(99:95, print(pop_song(drink = "beer")))

```

## You can kind of get around this...

```{r}

sapply(99:95, function(x) pop_song(x, drink = "beer"))

```

## But mapping has a shortcut!

```{r}

map_chr(99:95, ~pop_song(.x, "beer"))

```

- Note the "~" in front of the function - this is important!
- Note the "." before the "x" - this is important!


## What if I want to use many options?

```{r, echo = FALSE}

data <- data.frame(num = 99:95, drink = c("pop", "beer", "water", "lemonade","milk"))

```

```{r}
head(data)
```

## Use map2!

```{r}
map2_chr(data$num, data$drink, ~pop_song(.x, .y))
```

# Vectorized functions

## What does it mean to be vectorized?

```{r}
add_odds <- function(x,y){
  
  if(x %%2 == 1 & y %% 2 == 1){
    
    x + y
    
  } else {
    
   0
    
  }
  
}

```

## What does it mean to be vectorized?

```{r}

add_odds(1,3)

add_odds(2,4)


```

## What does it mean to be vectorized?

```{r, error = TRUE }

first_nums <- c(1,2)
next_nums <- c(3,4)

add_odds(first_nums, next_nums)

```

## What does it mean to be vectorized?

`add_odds` is not a vectorized function!

```{r, eval = FALSE}
add_odds <- function(x,y){
  
  if(x %%2 == 1 & y %% 2 == 1){
    
    x + y
    
  } else {
    
   0
    
  }
  
}

```

## What does it mean to be vectorized?

 We can try to make it vectorized...
 
```{r}
add_odds_2 <- function(x,y){
  
  is_odd <- x %% 2 == 1 & y %% 2 == 1
  
  answers <- rep(0, length(x))
  
  answers <- x + y
  
  answers[!is_odd] <- 0
  
}

```

## What does it mean to be vectorized?

```{r}

first_nums <- c(1,2)
next_nums <- c(3,4)

add_odds_2(first_nums, next_nums)

```

## But, what if we can't vectorize?

```{r}

first_nums <- c(1,2)
next_nums <- c(3,4)

map2(first_nums, next_nums, ~add_odds(.x, .y))

```

# Absolute references

## External objects

- A function you write can rely on outside information.

```{r}
a <- 5

add_to_a <- function(b){
  a + b
}

add_to_a(3)
```

## External objects

- But try to avoid this!  What if your information changes?

```{r, error = TRUE}
a <- NA

add_to_a(3)
```

## External objects

- But try to avoid this!  What if you don't have the object?

```{r, error = TRUE}
add_to_c <- function(c){
  a + c
}

add_to_c(3)
```




# Things to remember


## Things to remember

- What does your function take as INPUT?
- What does your function give as OUTPUT?
- How can we write it to be friendly with `apply` or `map`?
- Do you REALLY need to use `apply` or `map`, or is it vectorized?
- Bonus: How can we write it to be friendly with `%>%`?






