---
title: "Modeling and Statistics"
author: Kelly Bodwin
date: 
output: 
  ioslides_presentation:
    fontsize: 10pt
    incremental: true
  # beamer_presentation:
  #   theme: "boxes"
  #   colortheme: "beaver"
  #   fonttheme: "structurebold"
---

# Is this a statistics class or not???

```{r, include = FALSE}
titanic <- read.csv("https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Datasets/titanic.csv")
library(tidyverse)
library(lubridate)
library(broom)
```

# Sampling from data

## Our best friend the titanic dataset

```{r}
head(titanic)
```

## Sampling from data

```{r}
?sample

sample(10)

sample(titanic$Name, 10)

rand <- sample(1:nrow(titanic), 10)
titanic[rand,]

titanic %>%
  sample_n(10)

```




# Probability distributions

## Probability distributions

Recall from your previous statistics classes...

- A **random variable** is a value we don't know until we take a **sample**.
    + Coin flip:  Could be heads (0) or tails (1)
    + Person's height:  Could be anything from 0 feet to 20 feet.
- The possible values of a random variable, and how likely they are, is determined by the **distribution**.
    + Coin flip has 50% chance of heads, 50% chance of tails
    + Likelihood of heights follow a bell curve around 5 foot 7.
- Some distributions have names.  For this class, you need to know:
    + Uniform
    + Normal
    + t-distribution
    + Binomial
    + Chi-Square
    
## Distributions in R

- r for **random sampling**: generate a random value from the
distribution

```{r}
runif(3)
runif(3, min = 10, max = 20)

rnorm(3)
rnorm(3, mean = -100, sd = 50)

rt(3, df = 11)

rbinom(3, size = 10, prob = 0.7)

rchisq(3, df = 11)
```

## Distributions in R

- p for **probability**: cumulative distribution function, i.e.
compute P(X < x)

```{r}
pnorm(1.5)
pnorm(70, mean = 67, sd = 3)

1 - pnorm(70, mean = 67, sd = 3)
```

- What is the probability that a Chi-Square distribution with 35 degrees of freedom gives a value of 25 or less?
- What is the probability that 10 coin tosses give 6 or more heads?

## Distributions in R

- q for **quantile**: given a probability p, compute the x-value
such that P(X < x) = p.

```{r}
qnorm(.95)
qnorm(.95, mean = 67, sd = 3)
```

- If I randomly sample a number between 16 and 22, I can be 80% sure my number will be less than what?
- What is the critical value for a 1-sided t-test at the 0.05 significance level, with a sample size of 50?


## Distributions in R

- d for **density**: compute the height of the mass or density
curve/function
    + Continuous distributions: usually meaningless

- What is the probability of getting exactly 12 heads in 20 coin tosses?

# Let's practice!

## Let's practice!
- WITHOUT USING R, try to predict the following:

```{r}

punif(0.674)

pnorm(2)
qnorm(.975)

pchisq(0, df = 12)

qt(10, df = 16)

dbinom(1, size = 2, prob = .87)
pbinom(1, size = 2, prob = .87)

```

## Plotting your samples

```{r}
my_samples <- data.frame(x = rnorm(100))

ggplot(my_samples, aes(x)) + geom_histogram()

ggplot(my_samples, aes(x)) + 
  geom_histogram(bins = 10) 

ggplot(my_samples, aes(x)) + 
  geom_histogram(bins = 10, aes(y = ..density..))

ggplot(my_samples, aes(x)) + 
  geom_histogram(bins = 10, aes(y = ..density..)) + 
  stat_function(fun=dnorm, col = "red")
```


# Basic hypothesis tests in R


## t-tests

We would like to test if men or women paid more on the Titanic.

```{r}
?t.test

t.test(Fare ~ Sex, data = titanic)

my_test <- t.test(Fare ~ Sex, data = titanic)
tidy(my_test)  ## requires the 'broom' package!

my_test <- tidy(t.test(Fare ~ Sex, data = titanic))
my_test$p.value
1 - pt(my_test$statistic, df = nrow(titanic))
```

## Chisquared Tests

```{r}
?chisq.test

titanic %>% 
  count(Sex, Pclass) %>%
  spread(key = Sex, value = n) %>%
  select(female, male) %>%
  chisq.test() %>%
  tidy()
```


# Many models

## Multiple tests

```{r}

t_by_class <- titanic %>% 
  group_by(Pclass) %>% 
  nest()

my_tests <- map_df(t_by_class$data, 
                   ~tidy(t.test(Fare ~ Sex, data = .x)))

my_tests
```


