---
title: "Welcome to the Tidyverse"
output: pdf_document
fontsize: 14pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

\section{What is the Tidyverse?}

\includegraphics[width = 1in]{./Images/hex-tidyverse.png} \hspace{1cm}
\includegraphics[height = 1in]{./Images/data-science.png}

Because **R** is open-source, rather than a product released by a company such as SAS or Matlab, it relies heavily on user-created packages for its evolution.

The biggest revolution and modernization of **R** has come in the form of a collection of packages known as the **Tidyverse**.  The father of the Tidyverse, and author of its most famous packages, is Hadley Wickham.  However, many prominent statisticians and data scientists are now active contributors, including: Charlotte Wickham, Jenny Bryan, Yihui Xie, David Robinson, Julia Silge, and many many more.

This week, you learned "Base R": everything we did in today's activity did not require any extra packages.  From now on, you will be learning everything through the lens of the Tidyverse.


\section{Example}

Consider the following dataset:

```{r, echo = FALSE}
data(mtcars)
```

```{r}
str(mtcars)
```

Suppose we wish to accomplish the following edits to this data:
\begin{enumerate}
  \item Drop the variables `qsec` and `vs`, because we don't need them.
  \item Include only automatic cars (`am` = 1)
  \item Round the miles per gallon and weight of the cars to the nearest whole number.
  \item Define a "fast car" to be one with high gear, cylinders, or horsepower. Make a variable for this, and count up how many cars are fast.
\end{enumerate}

Here is how we would do it in Base R:

```{r}
data(mtcars)

mtcars <- mtcars[, -c(7,8)]

auto <- mtcars$am == 1
mtcars <- mtcars[auto, ]

mtcars$mpg <- round(mtcars$mpg)
mtcars$wt <- round(mtcars$wt)

mtcars$speed <- rep("slow", nrow(mtcars))

fast_cars <- (mtcars$gear == 5 | mtcars$cyl == 8 | mtcars$hp > 200)

mtcars$speed[fast_cars] <- "fast"

mtcars$speed <- factor(mtcars$speed)
summary(mtcars$speed)
```

\pagebreak
Here is the Tidy version:

```{r}
data(mtcars)

mtcars <- mtcars %>% 
  select(-qsec, -vs) %>%
  filter(am == 1) %>%
  mutate_at(
    vars(mpg, wt), 
    funs(round)
  ) %>%
  mutate(
    speed = case_when(
      gear == 5 | cyl == 8 | hp > 200 ~ "fast",
      TRUE ~ "slow"
    )
  )

mtcars %>% count(speed)
```


\section{Advantages}

The primary reasons to learn R via the Tidyverse, in my opinion:

\begin{itemize}
  \item {\bf Readability}: Code in the tidy style can be read more easily like an English sentence.
  \item {\bf Reproducibility}: When the flow is clearer, it is easier to build sequences that can be re-applied to new data.
  \item {\bf Presentation}: Tidy tools emphasize nice clean output. (RMarkdown is considered part of the Tidyverse!)
  \item {\bf Culture:} The guiding principles behind Tidyverse packages emphasize sharing of code and data, responsible use of statistics (reproducibility!), and collaboration.
  \item {\bf Community}: The Tidyverse R community is very active, diverse and welcoming. It is easy to get involved, and even the big names are happy to help beginners learn.
\end{itemize}


