---
title: "Lab Assignment 2: Avocado Prices"
output: pdf_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE,eval=FALSE)
```


# Instructions

Submit your .html file to the assignment on PolyLearn.

# Introduction

In this lab we're going to be looking at avocado prices! The dataset comes to us from kaggle and represents weekly retail scan data: avocado.csv.  A description of the data can be found at the [Hass Avocado Board website.](http://www.hassavocadoboard.com/retail/volume-and-price-data)

```{r, echo = TRUE}
library(tidyverse)
avo <- read.csv("https://www.dropbox.com/s/vsc1dkosz6nwake/avocado.csv?dl=1")
```

# Exercises


1) Which region sold the most bags of small organic avocados in 2017? 

*Hint: TotalUS does not count as a region!*

```{r}
avo %>% 
  filter(type == "organic", region != "TotalUS") %>%
  group_by(year, region) %>%
  summarize(tot_small_bags = sum(Small.Bags)) %>%
  filter(year == 2017) %>%
  distinct(region, .keep_all = TRUE) %>%
  top_n(1, tot_small_bags)
```

2) Use `separate()` to split the `Date` variable into year, month, and day.  In which month is the highest volume of avocado sales?

3) Which regions sell the most avocados by volume?  Plot side-by-side boxplots of Total Volume for only the 5 regions with the highest averages for the Total Volume variable. 

*Hint: Once you narrow down to the top 5 regions, you can use `pull()` to save the vector of region names for later use.*

```{r}
top5reg <- avo %>%
    group_by(region) %>%
    summarize(avgtotvol = mean(TotalVolume)) %>%
    arrange(desc(avgtotvol)) %>%
    slice(1:5) %>%
    pull(region)

avo %>%
    filter(region %in% top5reg) %>%
    ggplot(aes(x = region, y = TotalVolume, fill = region)) +
    geom_boxplot()
```


The following four California regions are in this dataset:  LosAngeles, SanDiego, Sacramento, SanFrancisco.  Answer the following questions about the California regions only.  

*Hint: These questions will require restructuring of your data!*

4) In which regions is the price of organic versus conventional avocados most different?  Support your answer with a few summary numbers and a plot.  

5)   How do their avocado habits differ?  Make a plot that shows, for all 4 California regions, the percent of avocado sales that are small, large, or extra large. Separate your plot by conventional vs. organic avocados.